---
title: "R Notebook"
output:
  html_notebook
---
Sources of this project is
- http://www2.uaem.mx/r-mirror/web/packages/copula/copula.pdf
- Enjoy the Joy of Copulas: With a Package copula
- https://stats.stackexchange.com/questions/184849/simulate-a-gaussian-copula-with-t-margins
Questions left:
- How do we consider std?
```{r}
library(MASS)
library(grid)
```
Two alternatives of sampling multivariate normal distribution 
```{r}
# Algorithm 3.2 in Quantitative Risk Managament
multi_norm <- function(){
  # Create matrix R
  rho <- sin((pi/2)*0.4) 
  num_of_z <- 50
  R <- matrix(rho,num_of_z, num_of_z)
  diag(R) <- 1
  # Cholesky decomposition
  chol_R <- t(chol(R))
  # Generate standard normal variables
  num_var <- dim(R)[1]
  z <- rnorm(num_var)
  # Multivariate normal distribution
  x <- chol_R %*% z
  return(x)
}
# https://www.rdocumentation.org/packages/MASS/versions/7.3-54/topics/mvrnorm
faster_multi_norm <- function() {
  # Create matrix R and vector mu
  rho <- sin((pi/2)*0.4) 
  num_of_z <- 50
  R <- matrix(rho,num_of_z, num_of_z)
  diag(R) <- 1
  mu <- matrix(0,num_of_z)
  # Sample
  x <- mvrnorm(mu = mu, Sigma = R)
  return(x)
}
```
Sampling multivariate Student's t dist.
```{r}
multi_t <- function() {
  # Create matrix R and vector mu
  rho <- sin((pi/2)*0.4) 
  num_of_z <- 50
  R <- matrix(rho,num_of_z, num_of_z)
  diag(R) <- 1
  mu <- matrix(0,num_of_z)
  # Consider example 9.4
  d.f <- 4
  # Normal dist.
  Z = mvrnorm(mu = mu, Sigma = R) # Returns RZ, control this
  # Chi dist.
  W = sqrt(d.f/rchisq(num_of_z,d.f))
  x <- mu + W  * Z
  return(x)
}
```

The copulas mentioned in the assignment and their sampling methods
```{r}
# Gaussian Copula
gaussian <- function(num_simu) {
  # Create a variable to store the res
  num_of_z <- 50 # 50
  res <- matrix(0,num_simu,num_of_z)
  # Declare constants for t-dist.
  std <- 0.01
  d.f <- 3
  for (i in 1:num_simu) {
    # Fetch the samples x_1, ..., x_d from N(0,R)
    x <- faster_multi_norm()
    # Calculate the cdf of each x_1, ..., x_d
    cdf_x <-pnorm(x)
    # Calculate the quantile w.r.t to normal dist.
    res[i,] <- std*qt(cdf_x,d.f)
  }
  return(res)
}

# t Copula
t <- function(num_simu) {
  # Create a variable to store the res
  num_of_z <- 50
  res <- matrix(0,num_simu,num_of_z)
  # Declare constants for t-dist.
  std <- 0.01
  d.f_4 <- 4
  d.f_3 <- 3
  for (i in 1:num_simu) {
    # Fetch the samples x_1, ..., x_d from N(0,R)
    x <- multi_t()
    # Calculate the cdf of each x_1, ..., x_d
    cdf_x <-pt(x, d.f_4)
    # Calculate the quantile w.r.t to normal dist.
    res[i,] <- std*qt(cdf_x,d.f_3)
  }
  return(res)
}

# Computes eq. 9.13 considering page 310 (before Eq. 9.14)
clayton.helper <- function () {
  num_of_z <- 50
  tau = 0.4
  theta = -(2*tau)/(tau-1)
  # Eq. 9.13
  X <- rgamma(1,1/theta)
  negative_log_V <- -log(runif(num_of_z))
  U <- matrix(0,num_of_z)
  for (i in 1:num_of_z) {U[i] <- psi(negative_log_V[i]/X, theta)}
  return(U)
}
psi <- function(u, theta) {
  return((u+1)^(-1/theta))
}
# Clayton Copula
clayton <- function(num_simu) {
  # Create a variable to store the res
  num_of_z <- 50 # 50
  res <- matrix(0,num_simu,num_of_z)
  # Declare constants
  std <- 0.01
  d.f <- 3
  for (i in 1:num_simu) {
    # Fetch the samples 
    x <- clayton.helper()
    # Calculate the quantile w.r.t to normal dist.
    res[i,] <- std*qt(x,d.f)
  }
  return(res)
}
```
The construction of log returns as given in page 308
```{r}
logreturn_func <- function(samples, simu){
  logreturn <- matrix(0,simu)
  V_0 <- 50*2*10^4
  for (i in 1:simu) {
    sample <- samples[i,]
    V_1 <- 2*10^4 * sum(exp(sample))
    logreturn[i] <- log(V_1/V_0)
  }
  return(logreturn)
}
```
The values of $Z_1, ..., Z_{50}$ given that they have a Gaussian copula.
```{r}
num_simu <- 100
z_c.g<- gaussian(num_simu)
logreturns_z_c.g <- logreturn_func(z_c.g, num_simu)
```
The empirical distribution function is given by $F_n(\vec{x}) = \frac{1}{n} \sum_{n=1}^n I(\vec{x}_k \leq \vec{x}).$ 
```{r}
emp_dist_func <- function(x, samples) {
  return (sum(samples < x) / length(samples)) 
}
emp_dist_plot <- function(samples) {
  samples <- sort(samples)
  len <- length(samples)
  interval <- samples[length(samples)] - samples[1]
  h <- len * length(sample)
  delta <- interval/h
  x <- matrix(0,h); y <- matrix(0,h)
  for (i in 1:h) {x[i] <- samples[1] + i * delta}
  for (i in 1:length(x)) {
    y[i] <- emp_dist_func(x[i],samples)
    }
  res <- matrix(0,2, length(x))
  rownames(res) <- c("x", "y")
  res[1,] <- x; res[2,] <- y;
  return(res)
}
```

The empirical distribution of $Z_1, ..., Z_{50}$ given that they have a Gaussian copula compare to the original values of $Z_1, ..., Z_{50}$
```{r, out.width='1\\linewidth', fig.asp=0.3, fig.ncol = 1, fig.cap="output",fig.align = "center"}
num_simu <- 500
# Gaussian Copula
z_c.g<- gaussian(num_simu)
# Student's t Copula
z_c.t <- t(num_simu)
# Clayton Copula
z_c.c <- clayton(num_simu)
logreturns_z_c.g <- logreturn_func(z_c.g, num_simu)
logreturns_z_c.t <- logreturn_func(z_c.t, num_simu)
logreturns_z_c.c <- logreturn_func(z_c.c, num_simu)

emp_dist_z_c.g <- emp_dist_plot(logreturns_z_c.g)
emp_dist_z_c.t <- emp_dist_plot(logreturns_z_c.t)
emp_dist_z_c.c <- emp_dist_plot(logreturns_z_c.c)
par(mfrow=c(1,3))
plot(emp_dist_z_c.g[1,], emp_dist_z_c.g[2,], cex = 0.3, col="black",
     ylab = "Empirical Distribution", xlab= "x", 
     main = "Gaussian Copula", panel.first = grid())
plot(emp_dist_z_c.t[1,], emp_dist_z_c.t[2,], cex = 0.3, col="black",
     ylab = "Empirical Distribution", xlab= "x", 
     main = "Student's t Copula", panel.first = grid())
plot(emp_dist_z_c.c[1,], emp_dist_z_c.c[2,], cex = 0.3, col="black",
     ylab = "Empirical Distribution", xlab= "x", 
     main = "Clayton Copula", panel.first = grid())
```
The empirical quantile distribution function is given by $F^{-1}_n(p) = x_[n(1-p)+1,n].$ 
```{r}
emp_quantile_func <- function(p, samples) {
  samples <- sort(samples, decreasing = TRUE)
  index <- floor(length(samples) *(1 - p)) + 1
  return (samples[index]) 
}
qq_plot <- function(samples, df = 3) {
  n <- length(samples)
  t_quan <- emp_quan <- matrix(0,n)
  for (k in 1:n) {
    index <- (n-k+1)/(n+1)
    t_quan[k]<- qt(index, df)
    emp_quan[k] <- emp_quantile_func(index, samples)
  }
  res <- matrix(0,2,length(t_quan))
  rownames(res) <- c("Reference", "Empirical")
  res[1,] <- t_quan; res[2,] <- emp_quan;
  return(res)
}

```

The empirical quantile function of $Z_1, ..., Z_{50}$ given that they have a Gaussian copula compared to the original values of $Z_1, ..., Z_{50}$.
```{r, out.width='1\\linewidth', fig.asp=0.3, fig.ncol = 1, fig.cap="output",fig.align = "center"}
num_simu <- 500
# Gaussian Copula
z_c.g<- gaussian(num_simu)
# Student's t Copula
z_c.t <- t(num_simu)
# Clayton Copula
z_c.c <- clayton(num_simu)
logreturns_z_c.g <- logreturn_func(z_c.g, num_simu)
logreturns_z_c.t <- logreturn_func(z_c.t, num_simu)
logreturns_z_c.c <- logreturn_func(z_c.c, num_simu)
qq_z_c.g <- qq_plot(logreturns_z_c.g, 3)
qq_z_c.t <- qq_plot(logreturns_z_c.t, 3)
qq_z_c.c <- qq_plot(logreturns_z_c.c, 3)
par(mfrow=c(1,3))
plot(qq_z_c.g[1,], qq_z_c.g[2,], cex = 0.3, col="black",
     ylab = "Empirical Quantile Distribution", xlab= "Student's t distribution", 
     main = "Gaussian Copula", panel.first = grid())
plot(qq_z_c.t[1,], qq_z_c.t[2,], cex = 0.3, col="black",
     ylab = "Empirical Quantile Distribution", xlab= "Student's t distribution", 
     main = "Student's t Copula", panel.first = grid())
plot(qq_z_c.c[1,], qq_z_c.c[2,], cex = 0.3, col="black",
     ylab = "Empirical Quantile Distribution", xlab= "Student's t distribution", 
     main = "Clayton Copula", panel.first = grid())

```

```{r}
V_1_func <- function(sample, sim) {
  # Store V_1 for each simulation
  V_1<- 2*10^4 * sum(exp(sample))
  return(V_1)
}
V_0_func <- function() {
  return(50*2*10^4)
}
var_func <- function(payoff) {
  R_0 <- 1
  L <- -1*payoff/R_0
  L <- sort(L, decreasing = TRUE)
  p <- 0.01
  index <- floor(length(L) * p) + 1
  return(L[index])
}
es_func <- function(payoff){
  R_0 <- 1
  L <- -1*payoff/R_0
  L <- sort(L, decreasing = TRUE)
  n <- length(L)
  p <- 0.01
  if (floor(n * p) == 0) stop("floor(n * p) == 0 is True ")
  index <- floor(n * p)
  es <- matrix(0,index)
  for (k in 1:index) {
    es[k] = (1/p)*(L[k]/n + ( p - index/n)*L[index + 1])
  }
  return(sum(es))
}
var_and_es_func <- function(samples, sim) {
  # Calculate the net payoff
  payoff <- matrix(0, sim)
  for (i in 1:sim) {
    sample <- samples[i,]
    payoff[i] <- V_1_func(sample) - V_0_func()
  }
  # Calculate VaR and ES
  var <- var_func(payoff)
  es <- es_func(payoff)
  res <- c(var, es)
  return(res)
}
```
Gaussian simulation
```{r, out.width='1\\linewidth', fig.asp=0.3, fig.ncol = 1, fig.cap="output",fig.align = "center"}
sim_arr <- seq(10^2,10^4,10^2) 
len_sim_arr <- length(sim_arr)
var <- matrix(0,len_sim_arr); es <- matrix(0,len_sim_arr);
table <- matrix(0,len_sim_arr,3); colnames(table) <- c("sim", "VaR", "ES")
for (i in 1:length(sim_arr)) {
  sample <- gaussian(sim_arr[i])
  res <- var_and_es_func(sample, sim_arr[i])
  var[i] <- res[1]
  es[i] <- res[2]
  table[i,1] <- sim_arr[i]; table[i,2] <- res[1]; table[i,3] <- res[2]
}
#print(table)

par(mfrow=c(1,2))
plot(var, cex = 0.3, col="black",
     ylab = "Value-at-Risk", xlab= "x", 
     main = "Gaussian Copula & Value-at-Risk", panel.first = grid())
plot(es, cex = 0.3, col="black",
     ylab = "Expected Shortfall", xlab= "x", 
     main = "Gaussian Copula & Expected Shortfall", panel.first = grid())
```
Student's simulation
```{r, out.width='1\\linewidth', fig.asp=0.3, fig.ncol = 1, fig.cap="output",fig.align = "center"}
sim_arr <- seq(10^2,10^4,10^2) 
len_sim_arr <- length(sim_arr)
var <- matrix(0,len_sim_arr); es <- matrix(0,len_sim_arr);
table <- matrix(0,len_sim_arr,3); colnames(table) <- c("sim", "VaR", "ES")
for (i in 1:length(sim_arr)) {
  sample <- t(sim_arr[i])
  res <- var_and_es_func(sample, sim_arr[i])
  var[i] <- res[1]
  es[i] <- res[2]
  table[i,1] <- sim_arr[i]; table[i,2] <- res[1]; table[i,3] <- res[2]
}
#print(table)

par(mfrow=c(1,2))
plot(var, cex = 0.3, col="black",
     ylab = "Value-at-Risk", xlab= "x", 
     main = "t Copula & Value-at-Risk", panel.first = grid())
plot(es, cex = 0.3, col="black",
     ylab = "Expected Shortfall", xlab= "x", 
     main = "t Copula & Expected Shortfall", panel.first = grid())
```
Clayton
```{r, out.width='1\\linewidth', fig.asp=0.3, fig.ncol = 1, fig.cap="output",fig.align = "center"}
sim_arr <- seq(10^2,10^4,10^2) 
len_sim_arr <- length(sim_arr)
var <- matrix(0,len_sim_arr); es <- matrix(0,len_sim_arr);
table <- matrix(0,len_sim_arr,3); colnames(table) <- c("sim", "VaR", "ES")
for (i in 1:length(sim_arr)) {
  sample <- clayton(sim_arr[i])
  res <- var_and_es_func(sample, sim_arr[i])
  var[i] <- res[1]
  es[i] <- res[2]
  table[i,1] <- sim_arr[i]; table[i,2] <- res[1]; table[i,3] <- res[2]
}

par(mfrow=c(1,2))
plot(var, cex = 0.3, col="black",
     ylab = "Value-at-Risk", xlab= "x", 
     main = "Clayton Copula & Value-at-Risk", panel.first = grid())
plot(es, cex = 0.3, col="black",
     ylab = "Expected Shortfall", xlab= "x", 
     main = "Clayton Copula & Expected Shortfall", panel.first = grid())
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


