---
title: "Risk Management - Project I"
output: html_notebook
---
packages necessary
```{r}

```
Consider a yearly investment of 1000 dollars in long positions in a portfolio of stocks
and a risk-free one-year zero coupon bond over a 30-year period. The yearly returns on
the portfolio of stocks year k is modeled as $Rk =\exp{\mu+\sigma Z_k}$ , where $Z_k\sim\mathcal{N}(0,1)$. The yearly returns are assumed to be independent. The yearly return on the risk-free bond is assumed to be $e^{r}$, where $r=0.01$. Each year, after the investment of 1000 dollars has been added, the entire portfolio is re-balanced such that the fraction of the total portfolio value invested in the stocks at the beginning of year $k$ is $p(1−c(k−1)/30)$, where $p, c \in [0, 1]$.

#######

Question(a) Determine a function f such that the value of the pension savings in 30 years can be expressed as $V_30 = f(\mu, \sigma, p, c,Z_1,...,Z_{30})$.

```{r}
portfolio_return <- function(mu,sigma,p,c,r,n=30){
  z <- rnorm(n)
  v <- 0*c(1:n)
  
  risky_investment = p*1000
  riskfree_investment = 1000 - risky_investment
  
  v[1] = risky_investment * exp(mu+sigma*z[1]) + riskfree_investment * exp(r)
  
  for(k in 2:n){
    risky_investment = p * (1 - c*(k-1)/30) * (v[k-1] + 1000)
    riskfree_investment = (v[k-1] + 1000) - risky_investment
    
    v[k] = risky_investment * exp(mu+sigma*z[k]) + riskfree_investment * exp(r)
  }
  return (v)
}
```

First look at $(p,c)$ and its corresponding function
```{r}
fraction <- function(p,c,k){
  f = p*(1-c*(k-1)/30)
  return (f)}
p = 0.5
c = 0.2
plot( ( function(k) p*(1-c*(k-1)/30)) ( c(1:30) ) , pch=18)
lines(( function(k) p*(1-c*(k-1)/30)) ( c(1:30) ))

```
Concluding: the $p$ is the start of the fraction invested in risky asset, which reduces to $p(\frac{30-29c}{30})$ in 30 years.

```{r}
# Initialization
# n = 30

mu = 0.1
sigma = 0.1
r = 0.01
```

Example of v-vector:
```{r}
#baseline
v <- portfolio_return(mu,sigma,p,c,r)
plot(1:30,v, col = 'red', pch='.')
lines(1:30,v, col = 'red',type="s")

#five differents portofolio returns
for (i in 1:9){
  v <- portfolio_return(mu,sigma,p,c,r)
  points(1:30, v, col = i, pch='.')
  lines(1:30, v, col = i, type="s")
}
```

Simulate a sample of suitable size n from the distribution of $(Z_1, . . . , Z_30)$ and use this sample to determine the empirical distribution $F_n$ of $V_30$ for a range of values of the pa- rameters $\mu, \sigma, p, c$.

```{r}
number_of_simulations = 1000
v_30_vector = 0*c(1:number_of_simulations) #matrix(ncol = number_of_simulations, nrow = 30)

#v_table = data.frame(Years = c(1:30))
#v_vector <- portfolio_return(mu,sigma,p,c,r)
#name <- paste("v(", 1,")",sep="") 
#df$name <- v_vector

for (j in 1:number_of_simulations){
  v_vector <- portfolio_return(mu,sigma,p,c,r)
  #name <- paste("v(", j,")",sep="") 
  #df$name <- v_vector
  v_30_vector[j] = v_vector[30]
}

hist(v_30_vector)
boxplot(v_30_vector,horizontal=TRUE,axes=TRUE,outline=TRUE)
Fn = ecdf(v_30_vector)
plot(Fn)
```

(b) Set $μ = 0.03$ and $σ = 0.2$ and investigate the effects on the empirical distribution $F_n(p,c)$ of $V_30$ of varying $p$ and $c$. Suggest a suitable criterion for selecting the optimal empirical distribution $F_n(p,c)$ and determine the optimizer $(p,c)$. For the optimal portfolio you must report the mean $E[V_30]$ and the 1%-quantile: $F^{−1}_{V_{30}}(0.01)$.

```{r}
expected_value = mean(v_30_vector)
expected_value

# fit gamma distr -> obtain fitted (alpha, beta) -> compute approximate expectation value E = alpha/beta
```

(c) Sweden’s pension system has transited from traditional life insurance to a system based on fund insurance. Roughly speaking fund insurance works as in (a) and (b) where the risky asset can be viewed as a portfolio selected by the customer. In traditional life in-surance the individual participates in a collective where there is a guaranteed return as well as a potential for higher returns for the collective if the investments in the risky asset (portfolio) performs well. One version of traditional life insurance works as follows (there are some variations on this theme).
There is a guaranteed annual return $\bar{r}$ on all investments. The guaranteed return is typically slightly lower than the risk-free rate. The fund manager of the traditional life insurance policy then invests the necessary amount in the risk-free asset to cover the guarantee and the remaining surplus in the risky asset to generate a higher return. For the investor with a yearly investment of 1000 dollars, the guaranteed amount after $k$ years is $G_k$ where $G_0 = 0$, $G_k=1000\sum_k \exp{(\bar{r}j)},k\geq 1$.
Let $V_0 = 0$ and $V_k, k \geq 1$, be the value of the portfolio after $k$ years. At the beginning of year $k + 1$ the amount $G_{k+1}\exp{−r}$ is invested in the bank account and $V_k + 1000 − G_{k+1}\exp{−r}$ is invested in the risky asset.

Let $\bar{r} = 0.005$ and plot a histogram of $V_30$ using the traditional life insurance strategy and compare it to your optimal solution in (b). You must also report the mean $E[V_30]$ and
the 1%-quantile: $F^{−1}(0.01)$.

```{r}
r_bar = 0.005
if( r_bar >= r) stop('risk-free rate is lower than the zero-coupon rate')

#G-vector
garanteed_return = c(1:30)
garanteed_return[1] = 1000 * exp(r_bar)
for(k in 2:30){
  garanteed_return[k] = garanteed_return[k-1] + 1000 * exp(k*r_bar)
}

#function for TradLiv
tradLiv_return <- function(mu,sigma,r, n = 30){
  z <- rnorm(n)
  v <- 0*c(1:n)
  
  riskfree_investment = garanteed_return[1] * exp(-r)
  risky_investment = 1000 - riskfree_investment
  
  v[1] = risky_investment * exp(mu+sigma*z[1]) + riskfree_investment * exp(r)
  
  for(k in 2:n){
    riskfree_investment = garanteed_return[k] * exp(-r)
    risky_investment = v[k-1] + 1000 - riskfree_investment
    
    if (risky_investment < 0) stop('Help were shortselling on the portfolio')
    
    v[k] = risky_investment * exp(mu+sigma*z[k]) + riskfree_investment * exp(r)
  }
  return (v)
}
```

Example of v-vector:
```{r}
v <- tradLiv_return(mu,sigma,r)
plot(1:30, v, col = 'purple', pch = '.')
lines(1:30,v, col = 'purple')

for (i in 1:10){
  v <- tradLiv_return(mu,sigma,r)
  points(1:30, v, col = i, pch = '.')
  lines(1:30,v, col = i)
}

for(i in 1:10){
v <- portfolio_return(mu,sigma,p,c,r)
points(1:30,v, col = 'blue',pch='.')
lines(1:30,v, col = 'blue', lwd =2)
}

v <- portfolio_return(mu,sigma,0,c,r)
points(1:30, v, col = 'red', pch = '.')
lines(1:30,v, col = 'red')
```

Run 1000 simulations
```{r}
number_of_simulations = 1000
v_30_vector = 0*c(1:number_of_simulations)

for (j in 1:number_of_simulations){
  v_vector <- tradLiv_return(mu,sigma,r)
  v_30_vector[j] = v_vector[30]
}

hist(v_30_vector)
boxplot(v_30_vector,horizontal=TRUE,axes=TRUE,outline=TRUE)
Fn = ecdf(v_30_vector)
plot(Fn)
```
d)
It may be difficult to generate large profits using the strategy in (c) and a way to obtain larger profits is to have a leverage in the exposure to stocks. In this case it is assumed that the market will not experience a crash of more than, say, 50% over one year.

At the beginning of year $k + 1$, the surplus over the guarantee is invested in the stock with a leverage of 2 (the 2 is $1/0.5$). That is, the amount $2(Vk + 1000 − G_{k+1}\exp{(−r)})$ is invested in the stock and the remaining amount, $V_k +1000−2(Vk + 1000 − G_{k+1}\exp{(−r)})$ in the bond. This type of strategy is often referred to as a constant proportion portfolio insurance (CPPI).
Note that if the stock price falls by more than 50% during year $k + 1$ the value $V_{k+1}$ is less than the guarantee $G_{k+1}$. In practice one would monitor the stock price continuously and rebalance to make sure that this ”never” happens. For this project you may assume that if you fall behind the guarantee, that is $V_{k+1} < G_{k+1}, then you invest the future yearly investments in the bond until Vk+m ≥ Gk+m and then proceed according to the strategy, or until the end (whatever comes first).
Compare the histogram of $V_30$ to the results in (b) and (c). You must also report the mean $E[V_30]$ and the 1%-quantile: $F^{−1}_{V_{30}}(0.01)$.

```{r}
alpha = 2

cppi_return <- function(mu,sigma,r,alpha, n=30){
  z <- rnorm(n)
  v <- 0*c(1:n)
  
  risky_investment = alpha * (1000 - garanteed_return[1] * exp(-r))
  riskfree_investment = 1000 - risky_investment
  
  v[1] = risky_investment * exp(mu+sigma*z[1]) + riskfree_investment * exp(r)
  
  for(k in 2:n){
    risky_investment = alpha * (v[k-1] + 1000 - garanteed_return[k] * exp(-r))
    riskfree_investment = v[k-1] + 1000 - risky_investment
    
    v[k] = risky_investment * exp(mu+sigma*z[k]) + riskfree_investment * exp(r)
  }
  return (v)
}
```

Example of v-vector:
```{r}
#vector_v = 0*c(1:1000)
#for (i in 1:1000){
#  v <- cppi_return(mu,sigma,r,alpha)
#  vector_v[i] = v[30]
#}
#print(mean(vector_v))
#print(sd(vector_v))

v <- cppi_return(mu,sigma,r,alpha)
plot(1:30, v, col = 'green', pch = '.', ylim=c(0,120000))#, log="x")#, xlim = c(25,30), ylim=c(4000,50000+100000))
lines(1:30,v, col = 'green')

for (i in 1:10){
  v <- cppi_return(mu,sigma,r,alpha)
  points(1:30, v, col = 'green', pch = '.')
  lines(1:30,v, col = 'green')
}

for (i in 1:10){
  v <- tradLiv_return(mu,sigma,r)
  points(1:30, v, col = 'red', pch = '.')
  lines(1:30,v, col = 'red')
}

for (i in 1:10){
v <- portfolio_return(mu,sigma,p,c,r)
points(1:30,v, col = 'blue',pch='.')
lines(1:30,v, col = 'blue', lwd = 1)
}

v <- portfolio_return(mu,sigma,0,c,r)
points(1:30, v, col = 'purple', pch = '.')
lines(1:30,v, col = 'purple')
```

```{r}
vector_of_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_returns[i] = cppi_return(mu,sigma,r,alpha)[30] #n=30
}

hist(vector_of_returns)
#P = ecdf(vector_of_returns)
#plot(P)

```

Comparing different methods

```{r}
vector_of_portfolio_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_portfolio_returns[i] = portfolio_return(mu,sigma,p,c,r)[30] #n=30
}
vector_of_tradliv_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_tradliv_returns[i] = tradLiv_return(mu,sigma,r)[30] #n=30
}
vector_of_cppi_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_cppi_returns[i] = cppi_return(mu,sigma,r,alpha)[30] #n=30
}
df = data.frame(
  pf = vector_of_portfolio_returns, tradliv = vector_of_tradliv_returns, cppi = vector_of_cppi_returns
)

#hist(vector_of_portfolio_returns)
#P = ecdf(vector_of_returns)
#plot(P)

```
```{r}
# using package plotrix
#install.packages("plotrix")
require(plotrix)

for (mu in c(0.1,0.2,0.3,0.4,0.5)){
  for(sigma in c(0.1,0.2,0.3,0.4,0.5)) {
    vector_of_portfolio_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_portfolio_returns[i] = portfolio_return(mu,sigma,p,c,r)[30] #n=30
}
vector_of_tradliv_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_tradliv_returns[i] = tradLiv_return(mu,sigma,r)[30] #n=30
}
vector_of_cppi_returns = c(1:number_of_simulations)
for (i in 1:number_of_simulations){
  vector_of_cppi_returns[i] = cppi_return(mu,sigma,r,alpha)[30] #n=30
}
df = data.frame(
  pf = vector_of_portfolio_returns, tradliv = vector_of_tradliv_returns, cppi = vector_of_cppi_returns
)
    multhist(df, freq = TRUE)
  }
}
#print(names(df))
```

